---
title: "SMPE Project"
author: "Insaf Boukrouh & Othmane Nahyl"
date: "30 décembre 2017"
output: 
  html_document: 
    fig_height: 10
    fig_width: 10
    highlight: tango
---
## Loading and cleaning the datasets
```{r, message=F, warning=F, cache=T}
library(plyr)
library(doParallel)
library(ggplot2)
library(grid)
library(gridExtra)

# Remove lines with NA values in the columns that we're interested in
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

# Register parallel backend in order to load files in parallel
nodes <- detectCores()
cl <- makeCluster(nodes)
registerDoParallel(cl)

# Load the csv files and keep certain columns
cities_data = ldply(list.files(path="../usa-data/",pattern="csv",full.names=TRUE),.parallel = TRUE,function(filename) {
  # Columns to keep
    keeps <- c("id","host_id","host_response_time","host_response_rate","host_acceptance_rate","host_total_listings_count","zipcode","country_code","property_type","room_type","accommodates","bathrooms","bedrooms","beds","bed_type","square_feet","price","security_deposit","cleaning_fee","extra_people","minimum_nights","maximum_nights","number_of_reviews","review_scores_rating","review_scores_accuracy","review_scores_cleanliness","review_scores_checkin","review_scores_communication","review_scores_location","review_scores_value","instant_bookable","cancellation_policy","host_since","calendar_last_scraped","city","state")
    city_data = read.csv(filename)
    city_data <- city_data[keeps]
    
    return(city_data)
})

stopCluster(cl)
```

## Cleaning the datasets
```{r, message=F, warning=F, cache=T}
# Convert host_response_rate from factors to numeric values
cities_data$host_response_rate <- as.numeric(gsub("%", "", as.character(cities_data$host_response_rate)))/100

# Convert host_acceptance_rate from factors to numeric values
cities_data$host_acceptance_rate <- as.numeric(gsub("%", "", as.character(cities_data$host_acceptance_rate)))/100

# Convert price from factors to numeric values
cities_data$price <- as.double(substring(gsub(",", "", as.character(cities_data$price)),2))

# Convert security_deposit from factors to numeric values
cities_data$security_deposit <- as.double(substring(gsub(",", "", as.character(cities_data$security_deposit)),2))
cities_data$security_deposit[is.na(cities_data$security_deposit)] <- 0

# Convert cleaning_fee from factors to numeric values
cities_data$cleaning_fee <- as.double(substring(gsub(",", "", as.character(cities_data$cleaning_fee)),2))
cities_data$cleaning_fee[is.na(cities_data$cleaning_fee)] <- 0

# Convert extra_people from factors to numeric values
cities_data$extra_people <- as.double(substring(gsub(",", "", as.character(cities_data$extra_people)),2))
cities_data$extra_people[is.na(cities_data$extra_people)] <- 0
```

## Different plots
```{r, message=F, warning=F}
summary(cities_data)
```

## Price plots according to different parameters
```{r, message=F, warning=F, cache=T}
cities_data_without_null_host_response_time <-subset(cities_data,host_response_time != "N/A" & host_response_time != "")
ggplot(data = cities_data_without_null_host_response_time, aes(x = host_response_time, y = price,color=host_response_time)) + geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1)+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(data = cities_data, aes(x = host_response_rate, y = price, color=host_response_rate)) + geom_point(size=0.1)

ggplot(data = cities_data, aes(x = host_total_listings_count, y = price, color=host_total_listings_count)) +geom_point(size=0.1) 

cities_data_without_null_property_type <-subset(cities_data,property_type != "N/A" & property_type != "")
ggplot(data = cities_data_without_null_property_type, aes(x = property_type, y = price,color=property_type))+ geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1)+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = cities_data, aes(x = room_type, y = price,color=room_type))+ geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = cities_data, aes(x = accommodates, y = price,color=accommodates)) +geom_jitter(position=position_jitter(0.3),size=0.1) 

ggplot(data = cities_data, aes(x = bathrooms, y = price, color=bathrooms)) +geom_jitter(width = 0.1,height = 0.2,size=0.1) 

ggplot(data = cities_data, aes(x = bedrooms, y = price,color=bedrooms)) +geom_jitter(position=position_jitter(0.3),size=0.1) 

ggplot(data = cities_data, aes(x = beds, y = price,color=beds)) +geom_jitter(position=position_jitter(0.3),size=0.1)

ggplot(data = cities_data, aes(x = bed_type, y = price,color=bed_type))+ geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1)

ggplot(data = cities_data, aes(x = square_feet, y = price,color=square_feet)) +geom_point(size=0.1)

ggplot(data = cities_data, aes(x = minimum_nights, y = price,color=minimum_nights))  +geom_point(size=0.1)

ggplot(data = cities_data, aes(x = maximum_nights, y = price,color=maximum_nights)) +geom_point(size=0.1)

ggplot(data = cities_data, aes(x = instant_bookable, y = price,color=instant_bookable))+ geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1)

ggplot(data = cities_data, aes(x = cancellation_policy, y = price,color=cancellation_policy))+ geom_boxplot() +geom_jitter(position=position_jitter(0.3),size=0.1) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
The plots show no correlations between the price and the majority of the parameters: the host response time, the host response rate, the host total listings count, the number of accommodates, the number of bathrooms, the number of bedrooms, the number of beds, the number of minimum and maximum nights (most of the listings don't have restrictions about these two parameters), whether the listing is instant bookable or not, and the cancellation policy (we can only see that the prices are slightly more expensive for the listings that have a strict cancellation policy).

For the square_feet, the great majority of the listings don't have this information. Thus, we can't conclude anything regarding it.

There are possible correlations between the price and:
- The property type: the prices are very expensive for some types like the earth houses and the tree houses.
- The room type: the entire homes are generally more expensive than the other types, and the shared rooms are the cheapest ones.
- The bed type: beds other than real ones are generally cheaper, but we have a big variability for the real beds.

In terms of variability, it's high for most of the boxplots. We also see that most of them are skewed towards the bottom, which indicates that the higher the prices are, the more spread out they are.

## Miscellaneous
```{r}
# function: removing lines with NA values in specific columns
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

# function: removing spaces
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

### Incomes
```{r, message=F, warning=F, cache=T}
# Clean
incomes_data <- completeFun(cities_data, c("price", "minimum_nights", "number_of_reviews", "host_since", "calendar_last_scraped"))

# Calculate the income for each listing
incomes_data$income <- incomes_data$number_of_reviews*incomes_data$minimum_nights*incomes_data$price

# This duration should be the same for listings having the same host 
incomes_data$duration <- abs(as.Date(as.character(incomes_data$calendar_last_scraped), format = "%Y-%m-%d") -
                          as.Date(as.character(incomes_data$host_since), format = "%Y-%m-%d"))/30

# Aggregate the income by host
incomes_data <- aggregate(
    cbind(incomes_data$income, incomes_data$duration),
    by = list(host_id = incomes_data$host_id),
    FUN = function(x) sum(x))

# Calculate the income per month
incomes_data$income_per_month = unlist(incomes_data['V1'])/unlist(incomes_data['V2'])

# Plot
ggplot(incomes_data, aes(x = income_per_month))+
  geom_histogram(color = "darkblue", fill = "lightblue", binwidth = 10)+
  labs(title = "Incomes in the USA", x = "Income per month", y = "Count")+
  coord_cartesian(xlim=c(0,500))
```
While a lot of people gain big amounts of money through AirBnB, most of them have an average income per month lower than 100 dollars.


### Type of listings
```{r, message=F, warning=F, cache=T}
# Get the room types and their percentages
room_types_counts <- table(cities_data$room_type)
room_types <- names(room_types_counts)
counts <- as.vector(room_types_counts)
percentages <- scales::percent(round(counts/sum(counts), 2))
room_types_percentages <- sprintf("%s (%s)", room_types, percentages)
room_types_counts_df <- data.frame(group = room_types, value = counts)

# Plot
pie <- ggplot(room_types_counts_df, aes(x = "", y = value, fill = room_types_percentages))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y", start = 0)+
  scale_fill_brewer("Room Types", palette = "Dark2")+
  ggtitle("Type of listings")+
  ylab("")+
  xlab("")+
  labs(fill="")+
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank())+
  geom_text(aes(label = percentages), size = 5, position = position_stack(vjust = 0.5))
pie
```
Most of the listings are entire homes or appartments, while the shared rooms don't represent more than 3% of the entire listings.

### Evolution of new hosts over time
```{r, message=F, warning=F, cache=T}
# Clean
new_hosts_data <- completeFun(cities_data, c("host_since"))

# Calculate the number of new hosts for each year (except for 2017)
new_hosts_data$host_since <- as.Date(new_hosts_data$host_since, '%Y-%m-%d')
new_hosts_data <- new_hosts_data[new_hosts_data$host_since < as.Date("2017-01-01"),]
new_hosts_data <- new_hosts_data[order(as.Date(new_hosts_data$host_since, format="%Y-%m-%d")),]
new_hosts_data$host_since <- format(as.Date(new_hosts_data$host_since, "%Y-%m-%d"), format="%Y-%m")
new_hosts_data_table <- table(new_hosts_data$host_since)

# Plot
plot(as.Date(paste(format(names(new_hosts_data_table), format="%Y-%m"),"-01", sep="")), as.vector(new_hosts_data_table), type = "l", xlab = "Time", ylab = "Number of new hosts")
```
The number of new hosts was increasing since 2008. However, there was a decrease of this number in the last two years.

### Airbnb Prices
```{r, message=F, warning=F, cache=T}
# Clean
prices_per_state <- completeFun(cities_data, c("price","state"))
prices_per_state$state <- trim(prices_per_state$state)
prices_per_state <- subset(prices_per_state, prices_per_state$state != '')
prices_per_state$state[prices_per_state$state == 'New York'] <- 'NY'
prices_per_state$state[prices_per_state$state == 'ny'] <- 'NY'
prices_per_state$state[prices_per_state$state == 'Baja California'] <- 'CA'
prices_per_state$state[prices_per_state$state == 'secc Terrazas'] <- 'CA'
prices_per_state$state[prices_per_state$state == 'ca'] <- 'CA'
prices_per_state$state[prices_per_state$state == 'Ca'] <- 'CA'
prices_per_state$state[prices_per_state$state == 'wa'] <- 'WA'
prices_per_state$state[prices_per_state$state == 'il'] <- 'IL'
prices_per_state$state[prices_per_state$state == 'Il'] <- 'IL'

# Calculate the average price per city
average_prices_per_state <- aggregate(cbind(prices_per_state$price),
                  by = list(state = prices_per_state$state),
                  FUN = function(x) mean(x))

# Plot
ggplot(data = average_prices_per_state, aes(x = average_prices_per_state$state, y = average_prices_per_state$V1))+
    geom_bar(stat = "identity", fill = "steelblue", width = 0.7)+
    geom_text(aes(label = round(average_prices_per_state$V1, 2)), size=4)+
    coord_flip()+
    xlab("State")+
    ylab("Average Daily Price")+  
    theme_minimal()
```
The average daily price of listings in the state of Vermont is very high (650 dollars), while it's the lowest in Northern Mariana Islands (90 dollars). However, we can't conclude anything from the plot since the average prices alone aren't representative of the whole prices in every state.

To make a conclusion, we have to analyze the variability of the prices.

```{r}
ggplot(data = prices_per_state, aes(x = state, y = price, color = state)) + geom_boxplot() + geom_jitter(position=position_jitter(0.3),size=0.1) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
The plot shows that we have a high variability of the prices in the majority of the states. In Vermont and the Northern Mariana Islands, the boxplots are almost flat. 

```{r}
vt_data = prices_per_state[prices_per_state$state == 'VT',]
nrow(vt_data)
mp_data = prices_per_state[prices_per_state$state == 'MP',]
nrow(mp_data)
```
Indeed, we have only one record from each of these two states. Thus, the plot of average daily prices was misleading.

### Regression to explain the Ratings
```{r, message=F, warning=F, cache=T}
scores <- c("review_scores_rating","review_scores_accuracy","review_scores_cleanliness","review_scores_checkin","review_scores_communication","review_scores_location","review_scores_value")
scores_data <- cities_data[scores]
scores_data <- completeFun(scores_data, scores)

ggplot(data = scores_data, aes(x = review_scores_accuracy, y = review_scores_rating, color=review_scores_accuracy)) + geom_jitter(size = 0.1)

ggplot(data = scores_data, aes(x = review_scores_cleanliness, y = review_scores_rating, color=review_scores_cleanliness)) + geom_jitter(size = 0.1)

ggplot(data = scores_data, aes(x = review_scores_checkin, y = review_scores_rating, color=review_scores_checkin)) + geom_jitter(size = 0.1)

ggplot(data = scores_data, aes(x = review_scores_communication, y = review_scores_rating, color=review_scores_communication)) + geom_jitter(size = 0.1)

ggplot(data = scores_data, aes(x = review_scores_location, y = review_scores_rating, color=review_scores_location)) + geom_jitter(size = 0.1)

ggplot(data = scores_data, aes(x = review_scores_value, y = review_scores_rating, color=review_scores_value)) + geom_jitter(size = 0.1)
```
From the plots, we can see that most of the people who give the listings high ratings, give high scores for all the other types of scores. However, this is not enough to assume the existence of correlations between the global and the specific scores.

## Trying to fit a linear model to explain the Ratings
```{r, message=F, warning=F, cache=T}
# Linear model with all variables
r_reg1 <- lm(data=cities_data, review_scores_rating~review_scores_accuracy+review_scores_cleanliness+review_scores_checkin+review_scores_communication+review_scores_location+review_scores_value)
summary(r_reg1)
```

## Analysis of the results
Our model has:
- A fairly good Adjusted R-squared value.
- A low residual standard error.
- Small standard errors of the coefficients of explanatory variables.

However, this is not enough to say that our linear model is good. To make sure of that, let's check whether the linear regression hypothesis hold or not.

```{r}
plot(r_reg1)
```
- The "Residuals vs Fitted" plot indicates if residuals have non-linear patterns. If the residuals are spread out around a horizontal line without distinct patterns, that means that we don't have non-linear relationships, which is not the case for our model since it seems that the residuals have some kind of structure.
- The "Normal Q-Q" plot shows if residuals are normally distributed. Our residuals are not well lined on the straight dashed line except in the middle of the plot, which is not quite good.
- The "Scale-Location" plot lets us check the assumption of equal variance (homoscedasticity). Our line is not horizontal with randomly spread points, thus, our residuals are not homoscedastic.
- The "Residuals vs Leverage" plot helps us find influential cases. In fact, even though data has outliers, they might not be influential to determine a regression line. In our plot, we can barely see Cook's distance lines because all cases are well inside of them. i.e: if we exclude the "52474" case for example, the changes in the slope coefficients won't be important.

To conclude, our linear model doesn't well explain the ratings.

# Price linear regression
```{r}
# Candidates explanatory variables
variables <- c("host_response_time","host_total_listings_count","property_type","room_type","accommodates","bathrooms","bedrooms","beds","bed_type","security_deposit","cleaning_fee","extra_people","minimum_nights","maximum_nights","number_of_reviews","instant_bookable","cancellation_policy","zipcode")

# Clean the data
my_data <- completeFun(cities_data, variables)
nrow(my_data)

# Linear regression with all the variables
reg <- lm(data=my_data, price~host_response_time+host_total_listings_count+property_type+room_type+accommodates+bathrooms+bedrooms+beds+bed_type+security_deposit+cleaning_fee+extra_people+minimum_nights+maximum_nights+number_of_reviews+instant_bookable+cancellation_policy+zipcode+square_feet)
summary(reg)

# Try ANOVA with all variables
anv <- aov(data=my_data, price~host_response_time+host_total_listings_count+property_type+room_type+accommodates+bathrooms+bedrooms+beds+bed_type+security_deposit+cleaning_fee+extra_people+minimum_nights+maximum_nights+number_of_reviews+instant_bookable+cancellation_policy+zipcode+square_feet)
summary(anv)

# Linear regression with only the significant variables
reg <- lm(data=my_data, price~host_total_listings_count+property_type+room_type+accommodates+bathrooms+bedrooms+beds+bed_type+security_deposit+extra_people+zipcode+square_feet)
summary(reg)

plot(reg)

# TODO: some other model

# Keep only two digits from the zipcode
# my_data$zipcode = as.numeric(substring(my_data$zipcode,1,2))
```





